The abundance of Bullet Groups in $\Lambda$CDM

### 2014-03-06 paper/ms.tex
conclusions

\title{Bullet Groups as a test of $\Lambda$CDM}
\author{J. G. Fern\'andez-Trincado$^{1,2,3}$, J. E. Forero-Romero$^1$
  and T. Verdugo$^3$} 
\affil{$^1$ Departamento de F\'{i}sica, Universidad de los Andes,
  Cra. 1 No. 18A-10, Edificio Ip, Bogot\'a, Colombia\\ 
       $^2$ Institute Utinam, CNRS UMR6213, Universit\'e de
  Franche-Comt\'e, OSU THETA de Franche-Comt\'e-Bourgogne,
  Besan\c{c}on, France\\ 
       $^3$ Centro de Investigaciones de Astronom\'ia, AP 264,
  M\'erida 5101-A, Venezuela}

@@

The Bullet Cluster provided a new kind of observational evidence of
the existece of dark matter. Quantifying the displacement between dark
matter and the dominant baryonic component (hot X-ray emitting gas)
\citep{ForeroRomero2011} has been used to test the CDM paradigm itself
by quantifying the substructure velocity requiered to produce such
displacement \citep{Hayashi2006} to finally estimate the expected
abundance of such events in the Universe.

@@

We use halo catalogs constructed using the BDM algorithm. [...]

@@

We clearly expect such displacement in the case where the
sub-structure has already passed through the center of the host
halo. In this section we attempt an estimation for the DM-baryon
displacement statistics. We work under the following
hypothesis. First, we consider that systems with $|\mu|<0.9$ have a
baryonic displacement, $d_{2D}^{\prime}$, equal to zero. Second, we
consider that all systems with infalling velocities $\mu<-0.9$ and
large displacements $D_{\rm off}>0.6$ also have baryonic displacements
equal to zero. Third, for all the other cases we estimate the
displacement between the baryons and the dominand DM peak by the
distances between the dominant peak and the center of mass of the main
halo, $d_{2D}^{\prime}=D_{\rm off} R_{\rm vir}$, where $X_{\rm off}$
is the offset computed for each host halo in the original catalog.

This simplified modeldoes not take into account that there is a
fraction of halos with $\mu<-0.9$ and $D_{\rm off}<0.6$ for which the
collision has not started and should have $d_{2D}^{\prime}=0$. A detailed
modeling of this fraction requires the study of the complete merger
tree of the halo and sub-halo, a study beyond the scope of this
Letter. Instead we cautio the reader that the derived fraction of
halos with a displacement $<d_{2D}^{\prime}$ can be considered as an
upper limit.

The results for the integrated distributions for $>d_{2D}^{\prime}$
are shown in Figure \ref{fig:baryonic_displacements}.[..]

@@

We selected four snapshot of the simulation for four different
redshifts (z$=0$, z$=0.25$, z$=0.5$ and z$=1$) based in the
distribution of redshift in the Figure 1 by \citet{verdugo}, and  for
each sample in redshift we selected the host halo with circular
velocities greater than 300 kms$^{-1}$ and was split in two  principal
groups: The first group correspond to host halo with circular
velocities between 300 kms$^{-1}$ to 700 kms${^{-1}}$ with mass in the
range of $10^{12}$ M$_\odot{}$ to $10^{14}$ M$_\odot{}$ in the range
of mass of the Bullet Groups, and second group correspond to host halo
with circular velocities greater than 700 kms$^{-1}$ with mass
$\geq10^{14}$ M$_\odot{}$, in the range of mass of the Bullet
Clusters, in  the Table \ref{table1} is shown the two groups selected
in this work. For each group, we  classified the corresponding
substructures most massive and associated with the corresponding host
halo.  In  this work,
we estimate the expected distribution of displacements
($d_{real,(X,Y)}$) in the projection 2-D that can estimated by
observations, these displacements correspond to the separation between
the minimal potential of the host halo and the  minimal potential of
the substructure, both are dark matter distributions.

\section*{Acknowledgements}

>>>>

\title{The abundance of Bullet Groups in $\Lambda$CDM}
\author{J. G. Fern\'andez-Trincado$^{1,2,3}$, J. E. Forero-Romero$^1$
  and T. Verdugo$^3$} 
\affil{$^1$ Departamento de F\'{i}sica, Universidad de los Andes,
  Cra. 1 No. 18A-10, Edificio Ip, Bogot\'a, Colombia\\ 
       $^2$ Institute Utinam, CNRS UMR6213, Universit\'e de
  Franche-Comt\'e, OSU THETA de Franche-Comt\'e-Bourgogne,
  Besan\c{c}on, France\\ 
       $^3$ Centro de Investigaciones de Astronom\'ia, AP 264,
  M\'erida 5101-A, Venezuela}

@@

The Bullet Cluster provided a new kind of observational evidence of
the existece of dark matter. Quantifying the displacement between dark
matter and the dominant baryonic component (hot X-ray emitting gas)
\citep{ForeroRomero2010} has been used to test the CDM paradigm itself
by quantifying the substructure velocity requiered to produce such
displacement \citep{Hayashi2006} to finally estimate the expected
abundance of such events in the Universe.

@@

We use halo catalogs constructed using the BDM algorithm. [...] \citep{Riebe2013}.

@@

We clearly expect such displacement in the case where the
sub-structure has already passed through the center of the host
halo. In this section we attempt an estimation for the DM-baryon
displacement statistics. We work under the following
hypothesis. First, we consider that systems with $|\mu|<0.9$ have a
baryonic displacement, $d_{2D}^{\rm bar}$, equal to zero. Second, we
consider that all systems with infalling velocities $\mu<-0.9$ and
large displacements $D_{\rm off}>0.6$ also have baryonic displacements
equal to zero. Third, for all the other cases we estimate the
displacement between the baryons and the dominand DM peak by the
distances between the dominant peak and the center of mass of the main
halo, $d_{2D}^{\rm bar}=D_{\rm off} R_{\rm vir}$, where $X_{\rm off}$
is the offset computed for each host halo in the original catalog.

This simplified modeldoes not take into account that there is a
fraction of halos with $\mu<-0.9$ and $D_{\rm off}<0.6$ for which the
collision has not started and should have $d_{2D}^{\rm bar}=0$. A detailed
modeling of this fraction requires the study of the complete merger
tree of the halo and sub-halo, a study beyond the scope of this
Letter. Instead we cautio the reader that the derived fraction of
halos with a displacement $<d_{2D}^{\rm bar}$ can be considered as an
upper limit.

The results for the integrated distributions for $>d_{2D}^{\rm bar}$
are shown in Figure \ref{fig:baryonic_displacements}.[..]

@@

In this Letter we estimate fraction of galaxy groups
that can present observational features associated to a
bullet-like event. This is motivated by the recent observational
results of \citep{Gastaldello} where a system (\bullg) on the mass range
$1\times 10^{14}$\hMsun and velocity dispersion $650$\kms was reported
to feature a displacement between its baryonic (hot gas) and dark
matter components.

Our estimate is based on large N-body DM-only cosmological
simulation. We estimate the distribution of projected displacements
between the dominant DM clumps in two kinds of systems; groups with circular
velocities $300\kms<V_{\rm c}<700\kms$ and clusters with $V_{\rm
  c}>700$\kms. We report these results at four different redshifts
$z=0.0,0.25,0.5$ and $10$.

We find that a fraction of $XX\pm XX\%$ of the halos in the group
sample can present a displacement equal or larger than the observed displacement
for \bullg. We also present these results in terms of the ratio of the
circular velocities of the sub-halo to the host halo to find that the
system \bullg is found within the expected parameter space spanned by
the simulation.

Next we describe the geometry of the collision by the angle between
the position and velocity vectors of the sub-halo with respect to the
host. This allows us to separate three different instances: the
sub-halo crosses the virial radio of the host, the sub-halo goes
through the center of the host and the secondary
infall after reaching apogee.

From this analysis we estimate a the integrated distribution of the
displacements between the DM and the baryonic component. In this case
we find that a fraction of $XX \pm XX\%$ halos in the group sample
should present displacements equal or larger than those observed for
the Bullet-Group \bullg.

The results we present here can be used as a potential test of
$\Lambda CDM$. (Comments about Gaels results on multimodal groups).
